### 2023-03-16 outputs/output.qmd
Updated outputs folder to include Quarto document

---
title: "Well-being of Individuals from the United States Global Social Survey are Affected by Finances, Relationships, and Health"
author: "Justin Abando, Jin Di Liu, Marcin Jaczynski"
thanks: "Code and data are available at: https://github.com/jackchinski/happiness_report"
date: "`r Sys.time()`"
date-format: "D MMMM YYYY"
abstract: "DON'T FORGET THIS"
format: pdf
bibliography: references.bib
---

```{r}
#| include: false

knitr::opts_chunk$set(echo=TRUE)

###Workplace Setup###
library(tidyverse)
library(dplyr)
library(ggplot2)

```

# Introduction

# Data Source

# Data Gathering

# Results

```{r}
#| label: fig-satfincome
#| fig-cap: Visualization of...
#| echo: false

ggplot( 
  clean5, aes(
    x = new_income_range, fill = satfin)) + 
  geom_bar(
    position = "stack") + #Creating a stacked bar plot
  theme_minimal() + #Using a simple theme
  theme(axis.text.x=element_text( #Adjusting horizontal axis text to be easily readable
    angle = 60, 
    hjust = 1)) + 
  labs(x = "Income Ranges", #Labeling the axes and fill section
       y = "Frequency",
       fill = "Financial Satisfaction")

```

# Discussion

# Appendix